---
title: "PA1_template"
author: "CJ"
date: "April 20, 2016"
output: html_document
---

this is an attempt at using R markdown and knitr to produce an easy to read document containing R code and R output.  This is for the Coursera class: Reproducible Research

# Reproducible Research: Assignment 1
## Loading and Prep of the data

Loading the Data
```{r}
activitydata <- read.csv("activity.csv")
```

Converting the date column to a date

```{r}

activitydata$date <- as.Date(activitydata$date)

```

## What is the mean total number of steps taken per day?

get total steps per day

```{r}
stepsperday <- tapply(activitydata$steps, activitydata$date, sum)

```

build histogram

```{r}
hist(stepsperday, main = "Histogram of Steps Per Day", xlab = "steps taken per day", ylab = "freqency")

```

provide mean

```{r}

mean(stepsperday, na.rm = "TRUE")

```

provide median

```{r}

median(stepsperday, na.rm = "TRUE")

```

## What is the average daily activity pattern

Make a timeseries plot of the average steps taken per interval

```{r}

intervalmean <- tapply(activitydata$steps, activitydata$interval, mean, na.rm = "TRUE")

plot(row.names(intervalmean), intervalmean, type = "l", xlab = "Interval", ylab = "Average", main = "Average Number of Steps Taken per Interval")

```



In which interval does the maximum occur?

```{r}

names(which.max(intervalmean))

```


## Imputing Missing Values

Find the number of missing values in the dataset

```{r}

sum(is.na(activitydata))

```


Devise a strategy for filling in all of the missing values in the dataset

Using the interval average to fill in gaps

```{r}
meansteps <- aggregate(steps ~ interval, data = activitydata, FUN = mean)
NAFill <- numeric()
for (i in 1:nrow(activitydata)) {
  observation <- activitydata[i,]
  if(is.na(observation$steps)){
    steps <- subset(meansteps, interval == observation$interval)$steps
  } else {
    steps <- observation$steps
  }
  NAFill <- c(NAFill, steps)
}

```


Create a new dataset that is equal to the original but with filled in missing data.


```{r}

new_data <- activitydata
new_data$steps <- NAFill


```

Make a histogram of the total number of daily steps and report the mean and median total daily steps.

```{r}

stepsperdayFill <- tapply(new_data$steps, new_data$date, sum)

```

And the Histogram is:

```{r}

hist(stepsperdayFill, main = "Histogram of Steps Per Day With Filled Data", xlab = "steps taken per day", ylab = "freqency")


```



Report on Mean of the new set


```{r}

mean(stepsperdayFill)

```




Report on Median of the new set

```{r}

median(stepsperdayFill)

```


Filling the NA values has had no impact on the daily average but a slight impact on the median.

## Are there any differences in activity patterns between weekdays and weekends

Create a new factor variable in the dataset with two levels "weekday"" and "weekend"

```{r}

day <- weekdays(activitydata$date)
daylevel <- vector()
for(i in 1:nrow(activitydata)){
  if(day[i] == "Sunday"){
    daylevel[i] <- "weekend"
  } else if (day[i] == "Saturday"){
    daylevel[i] <- "weekend"
  } else{
    daylevel[i] <- "weekday"
  }
}

activitydata$daylevel <- daylevel
activitydata$daylevel <- factor(activitydata$daylevel)

stepsbyday <- aggregate(steps ~ interval + daylevel, data = activitydata, mean)
names(stepsbyday) <- c("interval", "daylevel", "steps")

```

Make a panel plot comparing weekday activity against weekend activity.

```{r}
library(lattice)
xyplot(steps ~ interval | daylevel, stepsbyday, type = "l", layout = c(1, 2), main = "Weekend vs Weekday Activity", xlab = "interval", ylab = "Steps")

```

